Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-23 alloscore_manuscript/alloscore-application-shorter.Rnw
next decision theory intro iteration

Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be interpreted as the value of $F$ as an input to
the decision making process under the realized outcome $y$ of $Y$.

>>>>

Decision theory investigates how decisions are made by formalizing a decision as the process of selecting an
\emph{action} $x$ from a specified set $\mathcal{X}$ of potential actions while taking into consideration the possible
consequences of $x$ under future states of the world. Let $Y$ be a quantifiable aspect of the future which will help to
determine the consequences of $x$. $Y$ is a random variable inasmuch as it is indeterminate when the decision is made.
We refer to a realization $y$ of $Y$ as an \emph{outcome}. When a decision maker assumes $Y$ to have a specified
distribution $F_Y$ and $Y$ is the only aspect of the future relevant to the selection of $x$, the process is referred to
as a decision under risk.  The selection of $x$ without a full specification of the distribution of $Y$ is referred to
as a decision under uncertainty. Clearly almost any real-world decision will be a decision under uncertainty.  Yet many
foundational results in decision theory were guided (as we will be) by the intuition that much can be learned about a
decision making process by building a ``small world'' approximation of the real-world (after \cite{savage1954foundations})
in which the beliefs of a decision
maker are equivalent to a fully specified distribution of a $Y$ that exhaustiveley quantifies all relevant aspects of
the future, converting a decision under uncertainty to a decison under risk.

Examples of actions with which this paper is concerned are levels of investment in measures designed to mitigate severe
disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of
all feasible levels of investment. An outcome $y$ determining the consequences of such an action might be the number of
individuals who eventually become sick and would benefit from the mitigation measure. A fundamental strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the (lack of) success of $x$. In our example, $x$ will succeed to the extent that it meets the
realized need $y$, so that greater unmet need will incur greater loss.

One tool a decision maker can use to anticipate the consequences of an action $x$ is a forecast
$F$ of $Y$. And within a ``small world'' where $Y$ exhaustively quantifies the future, the decision maker might choose that
$x$ which yields the lowest expected loss according to $F$. The loss of the action $x$ chosen according to $F$ can then, 
in turn, be interpreted as the value of $F$ as an
input to the decision making process under the realized outcome $y$ of $Y$.
